---
title: "The BrailleR package Example 2"
author: "A. Jonathan R. Godfrey"
bibliography: BrailleRPublications.bib
vignette: >
  %\VignetteIndexEntry{Example 2: Basic numerical summaries}
    %\VignetteEngine{knitr::rmarkdown}
output: knitr:::html_vignette
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(fig.width=7, fig.height=5, comment="")
library(BrailleR)
```


# The BrailleR package Example 2.

## Basic numerical summaries

The standard presentation of a summary of a data frame where each variable is given its own column is difficult for a screen reader user to read as the processing of information is done line by line. For example:

```{r UseSummary}
summary(airquality)
```

The VI() command actually calls the VI.data.frame() command. It then processes each variable one by one so that the results are printed  variable by variable instead of summary statistic by summary statistic. For example:
```{r UseVI}
VI(airquality)
```



## Important features

Note that in this case, the blind student could choose to present the summary of each variable as generated by the VI() command, or the output from the standard summary() command. There is no difference in the information that is ultimately presented in this case.


