---
title: Testing the VI.ggplot() within the BrailleR package"
author: "A. Jonathan R. Godfrey"
bibliography: BrailleRPublications.bib
vignette: >
  %\VignetteIndexEntry{qplot}
  %\VignetteEngine{knitr::rmarkdown}
output: knitr:::html_vignette
---


N.B.  the commands here are either exact copies of the commands presented in Wickham (2009) or some minor alterations to them.    Notably, some code given in the book no longer works. This is given a `#!`    


The `ggplot2` package has  a `summary` method that often but not always offers something to show that things have changed from one plot to another. I've kept the use of this command in as it shows where we need to look for more information and from where we ought to look for it.

```{r GetLibraries}     
library(BrailleR)   
library(ggplot2)   
library(magrittr)
dsmall = diamonds[1:100,]   
```  

```{r g1}     
g1 = qplot(carat, price, data = diamonds)   
summary(g1)   
g1  
VI(g1)  
```  

```{r g2}          
g2 = qplot(carat, price, data = dsmall, colour = color)    
summary(g2)   
g2
g2 %>% VI()   
# to get the content of the graph to be printed we need:   
#class(g2)="list"; g2   
```  

```{r g3}     
g3 = qplot(carat, price, data = dsmall, shape = cut)    
summary(g3)   
g3
VI(g3)   
#class(g3)="list"; g3   
```  

```{r g4}     
# to get semi-transparent points   
g4 = qplot(carat, price, data = diamonds, alpha = I(1/100))    
summary(g4)   
g4
VI(g4)   
```  

```{r g5}     
# to add a smoother (default is loess for n<1000)   
g5 = qplot(carat, price, data = dsmall, geom = c("point", "smooth"))    
summary(g5)   
g5
VI(g5)   
#! g5a = qplot(carat, price, data = dsmall, geom = c("point", "smooth"), span = 1)    
library(splines)    
#! g5b = qplot(carat, price, data = dsmall, geom = c("point", "smooth"), method = "lm")    
#! g5c = qplot(carat, price, data = dsmall, geom = c("point", "smooth"), method = "lm", formula = y ~ ns(x,5))    
```  

```{r g6}     
# continuous v categorical   
g6 = qplot(color, price / carat, data = diamonds, geom = "jitter", alpha = I(1 / 50))    
summary(g6)   
g6
VI(g6)   
g6a = qplot(color, price / carat, data = diamonds, geom = "boxplot")    
summary(g6a)   
g6a
VI(g6a)   
```  

```{r g7}     
# univariate plots   
g7a = qplot(carat, data = diamonds, geom = "histogram")    
summary(g7a)   
g7a
VI(g7a)   
g7b = qplot(carat, data = diamonds, geom = "histogram", binwidth = 1, xlim = c(0,3))    
g7c = qplot(carat, data = diamonds, geom = "histogram", binwidth = 0.1, xlim = c(0,3))    
g7d = qplot(carat, data = diamonds, geom = "histogram", binwidth = 0.01, xlim = c(0,3))    
summary(g7d)   
g7d
VI(g7d)   
```  

```{r g8}     
g8 = qplot(carat, data = diamonds, geom = "density")    
summary(g8)   
g8
VI(g8)   
```  

```{r g9}     
# data is separated by implication using the following...   
g9 = qplot(carat, data = diamonds, geom = "density", colour = color)    
summary(g9)   
g9
VI(g9)   
g10 = qplot(carat, data = diamonds, geom = "histogram", fill = color)    
summary(g10)   
g10
VI(g10)   
```  

```{r g11}     
# bar charts for categorical variable   
g11a = qplot(color, data = diamonds)    
summary(g11a)   
g11a
VI(g11a)   
g11b = qplot(color, data = diamonds, geom = "bar")    
summary(g11b)   
g11b
VI(g11b)   
g12a = qplot(color, data = diamonds, geom = "bar", weight = carat)   
summary(g12a)   
g12a
VI(g12a)   
g12b = qplot(color, data = diamonds, geom = "bar", weight = carat) + scale_y_continuous("carat")    
summary(g12b)   
g12b
VI(g12b)   
```  

```{r g13}     
# time series plots   
g13a = qplot(date, unemploy / pop, data = economics, geom = "line")    
summary(g13a)   
g13a
VI(g13a)   
g13b = qplot(date, uempmed, data = economics, geom = "line")    
summary(g13b)   
g13b
VI(g13b)   
```  

```{r g14}     
# path plots   
year <- function(x) as.POSIXlt(x)$year + 1900    
g14a = qplot(unemploy / pop, uempmed, data = economics, geom = c("point", "path"))    
summary(g14a)   
g14a
VI(g14a)   
#g14b = qplot(unemploy / pop, uempmed, data = economics, geom = "path", colour = year(date)) + scale_area()    
#summary(g14b)   
#VI(g14b)   
```  

```{r g15}     
# facets is the ggplot term for trellis' panels   
g15a = qplot(carat, data = diamonds, facets = color ~ ., geom = "histogram", binwidth = 0.1, xlim = c(0, 3))    
summary(g15a)   
g15a
VI(g15a)   
g15b = qplot(carat, ..density.., data = diamonds, facets = color ~ ., geom = "histogram", binwidth = 0.1, xlim = c(0, 3))   
summary(g15b)   
g15b
VI(g15b)   
```  

```{r g16}     
# rescaling of the axes   
g16 = qplot(carat, price, data = dsmall, log = "xy")   
summary(g16)   
g16
VI(g16)   
```  

```{r g17}     
g17 = qplot(displ, hwy, data=mpg, facets =~ year) + geom_smooth()    
summary(g17)   
g17
VI(g17)   
```  
